Komols Complexity, Coresets, and Sketches in Machine Learning

### 2018-12-26 komlos.tex
draft gammma2

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{conjecture}{Conjecture}[section]
%\newtheorem{claim}[theorem]{Claim}
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{observation}[theorem]{Observation}
\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}
%\newtheorem*{definition*}{Definition}
%\newtheorem{definition}{Definition}
%\newtheorem{claim}{Claim}
%\newtheorem{theorem}{Theorem}

@@

In \cite{DBLP:journals/corr/abs-1802-01751} (see paragraph before Lemma 2.2) they manage to obtain a coreset (with equal weights) of size $\sqrt{d \log(1/\eps)}/\eps$ for any kernel with certain properties. They do so by considering a matrix whose rows correspond to all the vectors in an $\eps$-net over the space, and columns correspond to the $n$ data points. Calling that matrix $A$, $A_{q,x} = K(q,x)$. They bound the discrepancy of the matrix defined as $\min_s \|As\|_\infty$ where $s$ is a sign vector by using a bound 
$$ \min_s \|As\|_\infty \leq \sqrt{\log(N)} \|A\|_{\gamma_2} $$
Here, $N$ is the size of the $\eps$ net and the $\gamma_2$ norm is 
$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \text{max norm of a row of B or C}$$
They prove that for psd kernel $K$, $\|A\|_{\gamma_2} \leq 1$ (in fact it is equal to 1). Now, back to our case. For any fixed power $k$ we know that $K(x,q)=\ip{x,q}^k$ is a psd kernel, hence its corresponding matrix $A_k$ has $\gamma_2$ norm bounded by $1$. Now, the logistic function can be written in a taylor expansion whose coefficients, in absolute value, sum to a constant $C$. It follows that the $A$ matrix for the logistic function has a $\gamma_2$ norm bounded by $C$. The same arguments as in  \cite{DBLP:journals/corr/abs-1802-01751} can now prove that its discrepancy is bounded by $\sqrt{\log(N)} \approx \sqrt{d\log(1/\eps)}$

@@

\section{Active Learning for linear functions (based on $\gamma_2$ norm)}

Let $x_1,\ldots,x_n \in R^d$ be and $y_1,\ldots,y_n$ be labels associated to the data points. Let $K$ be a psd kernel. Consider a problem of learning a linear function aiming to minimize the loss of 
$$\min_w \sum_{i=1}^n L(K(x_i, w), y_i)$$

@@

\begin{definition}
For a matrix $A$ of dimensions $m \times n$, $\disc(A)$ is defined as 
$$ \min_{s \in \{-1,1\}^n} \|As\|_\infty $$
\end{definition}
In \cite{matouvsek2014factorization} they prove that 
$$\|A\|_{\gamma_2} / \log(m) \leq \disc(A) \leq \|A\|_{\gamma_2}  \sqrt{\log(m)}$$
This gives a way to control the discrepancy of a matrix by representing it as a sum of matrices with known discrepency or $\gamma_2$ norm. In particular, for the logistic function we get that 
$$\disc(A_{f_L}) \leq \sqrt{\log(|Q|)} $$
where $Q$ is an $\eps$-net over the Euclidean unit sphere. This gives the bound of 
$$\disc(A_{f_L}) \leq O(\sqrt{d\log(1/\eps)}/\eps) $$

@@

\begin{corollary}
For $L(z)=\log(1+\exp(z))$ and $L(z) = z^2$, and any psd kernel $K:\R^d \times \R^d \to \R$ we have
$$\|A_{K,f_L}\|_{\gamma_2} \leq 1 $$ 
and
$$\disc(A_{K,f_L}) \leq O(\sqrt{d\log(1/\eps)}/\eps) $$ 
\end{corollary}

%Let's discuss what this means. For a logistic regression problem, by observing the weights we can change the signs of the inputs so that the logistic loss is always $\log(1+\exp(\ip{q,x}))$, and we can now convert the discrepancy result into a coreset of size $\sqrt{d\log(1/\eps)}/\eps$, as opposed to a random sample of $1/\eps^2$. This has a limitation where we need the labels to determine the coreset. We now proceed to see how we can remove that need.
%
%We will require an assumption regarding the labels of the data: We assume that for some $q \in Q$, there exist at most $\rho n$ data point label pairs $(x,y) \in X \times \{-1,1\}$ such that $\text{sign}(\ip{q,x}) \neq y$.

>>>>

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{conjecture}{Conjecture}[section]
%\newtheorem{claim}[theorem]{Claim}
%\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{observation}[theorem]{Observation}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}
%\newtheorem*{definition*}{Definition}
%\newtheorem{definition}{Definition}
%\newtheorem{claim}{Claim}
%\newtheorem{theorem}{Theorem}

@@

In \cite{DBLP:journals/corr/abs-1802-01751} (see paragraph before Lemma 2.2) they manage to obtain a coreset (with equal weights) of size $\sqrt{d \log(1/\eps)}/\eps$ for any kernel with certain properties. They do so by considering a matrix whose rows correspond to all the vectors in an $\eps$-net over the space, and columns correspond to the $n$ data points. Calling that matrix $A$, $A_{q,x} = K(q,x)$. They bound the discrepancy of the matrix defined as $\min_s \|As\|_\infty$ where $s$ is a sign vector by using a bound 
$$ \min_s \|As\|_\infty \leq \sqrt{\log(N)} \|A\|_{\gamma_2} $$
Here, $N$ is the size of the $\eps$ net and the $\gamma_2$ norm is 
$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \text{max norm of a row of B or C}$$
They use the fact that for psd kernel $K$, $\|A\|_{\gamma_2} \leq 1$ (in fact it is equal to 1). Now, back to our case. For any fixed power $k$ we know that $K(x,q)=\ip{x,q}^k$ is a psd kernel, hence its corresponding matrix $A_k$ has $\gamma_2$ norm bounded by $1$. Now, the logistic function can be written in a taylor expansion whose coefficients, in absolute value, sum to a constant $C$. It follows that the $A$ matrix for the logistic function has a $\gamma_2$ norm bounded by $C$. The same arguments as in  \cite{DBLP:journals/corr/abs-1802-01751} can now prove that its discrepancy is bounded by $\sqrt{\log(N)} \approx \sqrt{d\log(1/\eps)}$

@@

\section{Coreset for linear functions (based on $\gamma_2$ norm)}

Let $x_1,\ldots,x_n \in R^d$ be and $y_1,\ldots,y_n$ be labels associated to the data points. Let $K$ be a psd kernel. Consider a problem of learning a linear function aiming to minimize the loss of 
$$\min_q \sum_{i=1}^n L(K(x_i, q), y_i)$$

Our objective is to find a core-set of the $n$ data points. That is, we aim to find a set $S \subseteq [n]$ such that for every $q$ it holds that 
$$ \left| \sum_{i=1}^n L(K(x_i, q), y_i) - \frac{n}{|S|} \sum_{i \in S} L(K(x_i, q), y_i) \right| \leq \eps n $$

@@

\begin{definition}
For a matrix $A$ of dimensions $m \times n$, $\disc(A)$ is defined as 
$$ \min_{s \in \{-1,1\}^n} \|As\|_\infty $$
\end{definition}
In \cite{matouvsek2014factorization} they prove that 
$$\|A\|_{\gamma_2} / \log(m) \leq \disc(A) \leq \|A\|_{\gamma_2}  \sqrt{\log(m)}$$
This gives a way to control the discrepancy of a matrix by representing it as a sum of matrices with known discrepency or $\gamma_2$ norm. In particular, for the logistic function we get that 
$$\disc(A_{f_L}) \leq \sqrt{\log(|Q|)} $$
where $Q$ is an $\eps$-net over the Euclidean unit sphere. This gives the bound of 
$$\disc(A_{f_L}) \leq O(\sqrt{d\log(1/\eps)}) $$

@@

\begin{corollary}
For $L(z)=\log(1+\exp(z))$ and $L(z) = z^2$, and any psd kernel $K:\R^d \times \R^d \to \R$ we have
$$\|A_{K,f_L}\|_{\gamma_2} \leq 1 $$ 
and
$$\disc(A_{K,f_L}) \leq O(\sqrt{d\log(1/\eps)}) $$ 
\end{corollary}
